---
title: "Join Lab"
author: "Claudia Dominguez"
date: "DATA 2401"
output: "html_document"
---


# Joins and `dplyr` practice

Load the `"nycflights13"` package.  You'll also need to load `dplyr` or `tidyverse`, if you didn't above.

```{r}
library(tidyverse)
#install.packages("nycflights13")
library(nycflights13)
my_flights <- flights # store my own version
my_flights <- na.omit(my_flights) ## this dataset has data on cancelled flights too, 
                                  ## so we 'll remove those for now

```


Create a dataframe of the average arrival delays for each _destination_

```{r}
library(tidyverse)

avg_arrival_delays <- my_flights %>%
  group_by(dest) %>%
  summarise(Average_Arrival_Delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(Average_Arrival_Delay))  

print(avg_arrival_delays)


```

Then use `left_join()` to join on the "airports" dataframe, which has the airport information.

```{r}
library(tidyverse)
library(nycflights13)

avg_arrival_delays <- my_flights %>%
  group_by(dest) %>%
  summarise(Average_Arrival_Delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(Average_Arrival_Delay))

avg_arrival_delays_with_info <- avg_arrival_delays %>%
  left_join(airports, by = c("dest" = "faa"))

print(avg_arrival_delays_with_info)

```


Note here: the column names don't match anymore! 

Which airport had the largest average arrival delay?

```{r}
largest_delay_airport <- avg_arrival_delays_with_info %>%
  slice_max(Average_Arrival_Delay, n = 1)

print(largest_delay_airport)

```



What is the average departure delay for each airline? To answer this with something more than just the FAA code of the airline name, you need to create a dataframe of the average departure delay for each _airline_, then use `left_join()` to join on the "airlines" dataframe


```{r}
library(tidyverse)
library(nycflights13)

avg_departure_delays <- my_flights %>%
  group_by(carrier) %>%
  summarise(Average_Departure_Delay = mean(dep_delay, na.rm = TRUE)) %>%
  arrange(desc(Average_Departure_Delay))

avg_departure_delays_with_info <- avg_departure_delays %>%
  left_join(airlines, by = "carrier")

print(avg_departure_delays_with_info)


```


Which airline had the smallest average departure delay? 


```{r}
smallest_delay_airline <- avg_departure_delays_with_info %>%
  slice_min(Average_Departure_Delay, n = 1)

print(smallest_delay_airline)



```





Examples from R4DS:    

```{r}

top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest

```

Now you want to find each flight that went to one of those destinations. You could, of course, filter. But! Now you can semi-join: 

```{r}
flights %>%
  left_join(airports, by = c("dest" = "faa"))

```


What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?
Perform the two joins below and consider what the results are. 

```{r}
missing_airports_in_flights <- anti_join(flights, airports, by = c("dest" = "faa"))
print(missing_airports_in_flights)

unused_airports <- anti_join(airports, flights, by = c("faa" = "dest"))
print(unused_airports)

```